---
title: "Final Report"
author: "Nikki Johnson, Molly Hischke, Collin Brehmer, and Kyle Hancock"
date: "12/9/2019"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

## Introduction

The Daily Double is an important part of the game of Jeopardy. If a player can effectively predict where the Daily Doubles will be on the board they have a clear advantage to win the game. Recent Jeopardy contestant James Holzhauer became known for his strategy of hopping around the board to find the Daily Doubles and then betting larger than normal quantities. In just 32 games, his total winnings reached almost $2.5 million (an average of over $75,000 per game).

What positions on the Jeopardy board are most likely to be a Daily Double? Recently, several people have attempted to answer this question and their results can be found in a variety of different sources. Finding the answer to this question is not as simple as it seems at first. The common datasets for Jeopardy questions do not contain the questions that were not asked. This can make finding the exact x and y positions difficult. Multiple approaches can be taken to deal with this issue. The approach taken to develop the figures in this report are described in the methods section. The answer to this question can be visualized through a heatmap that shows each position on the Jeopardy board. Similar to previous analyses, the results in this report are presented in a heatmap to determine if there was a difference in daily double position by round. As a secondary question, this analysis faceted the results by year to evaluate changes over time, which has not been included in previous analyses. This important question presents a challenge in data cleaning, but when done correctly can result in interesting figures that have drawn the attention of many Jeopardy fans and data enthusiasts.

## Methods: 

The Jeopardy dataset utilized was found on GitHub which included the air date, round, daily double information, category, value, question, answer, comments, and notes. The dataset was filtered to only include dates beyond November 26, 2001. On that date, there was a change in the point values within the rounds (for example, round one went from values of 100, 200, 300, 400, and 500 to 200, 400, 600, 800 and 1000). During the point value update, there could have been other changes made to Jeopardy impacting the daily double locations (i.e. changes in how the positions were assigned). Therefore, we have excluded dates prior to the change. 

There were two main challenges in finding the positions of the daily double questions:

1. Finding the x position (the location of the categories).
2. Finding the y position (the location of the values). 

Not all categories (n = 12) were avaiaible in the dataset for all episodes. To solve this issue, only episodes where all 12 categories were played were included. The assumption was that the dataset listed the categories in order by x location within episode (i.e. first category to appear was assigned a x position of 1). 

In addition, not every question (n = 5) within a category was available in the dataset. This was especially challenging when a category contained a daily double question. Since players select their own bid for daily doubles, the value within the dataset gave little information about the location on the y axis. If there were 2 missing questions and 1 daily double within a category, the daily double could be at any of the missing value locations (Figure 1). To solve the issue, the daily double was weighted across the unasked questions in the category (i.e. 2 missing questions in a daily double category would result in a 1/2 daily double at each missing position). 

```{r, echo = FALSE}
library(tidyverse)
library(ggplot2)
jeopardy_board <- tibble(
  x_pos = c("Category 1", "Category 1", "Category 1", 
            "Category 1", "Category 1", 
            "Category 2", "Category 2", "Category 2", 
            "Category 2", "Category 2",
            "Category 3", "Category 3", "Category 3", 
            "Category 3", "Category 3", 
            "Category 4", "Category 4", "Category 4", 
            "Category 4", "Category 4", 
            "Category 5", "Category 5", "Category 5", 
            "Category 5", "Category 5",
            "Category 6", "Category 6", "Category 6", 
            "Category 6", "Category 6"),
  y_pos = c(200, 400, 600, 800, 1000, 
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000, 
            200, 400, 600, 800, 1000))

jeopardy_board <- jeopardy_board %>% 
  mutate(number_of_dd = c(1, 0, 0, 2, 1,
                          1, 0, 1, 1, 1,
                          1, 1, 1, 1, 1,
                          1, 1, 1, 0, 1,
                          1, 0, 0, 1, 1,
                          1, 1, 1, 0, 1),
          text = c(NA, "MISSING", "MISSING", "Daily Double", NA,
                          NA, "MISSING", NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, "MISSING", NA,
                          NA, "MISSING", "MISSING", NA, NA,
                          NA, NA, NA, "MISSING", NA)) 
weighted_jeopardy_board <- jeopardy_board %>% 
  mutate(number_of_dd = c(NA, 0.33, 0.33, 0.33, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA)) 

```


```{r, echo = FALSE, fig.height= 4, fig.width = 7}
jeopardy_board %>% 
   mutate(y_pos = factor(y_pos), 
                        y_pos = factor(y_pos, levels = rev(levels(y_pos)))) %>% 
  ggplot(aes(x = x_pos, y = y_pos, fill = number_of_dd)) +
  geom_tile(color = "White", size = 0.1) +
  scale_fill_continuous(high = "#132B43", low = "white") + 
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(legend.position = "none") + 
  geom_text(aes(label = text), color = "gray") +
  scale_x_discrete(position = "top") +
  labs(caption = paste0("Figure 1. In Category 1, the 400, 600 and 800 values",
  " were missing, and \na daily double question was asked. But it's unknown ",
  "if the daily double was \nat the 400, 600 or 800 value position.")) +
  theme(plot.caption = element_text(hjust = 0, size = rel(1.2)))
```

```{r, echo = FALSE, fig.height= 4, fig.width = 7}
weighted_jeopardy_board %>% 
   mutate(y_pos = factor(y_pos), 
                        y_pos = factor(y_pos, levels = rev(levels(y_pos)))) %>% 
  ggplot(aes(x = x_pos, y = y_pos, fill = number_of_dd)) +
  geom_tile(color = "White", size = 0.1) +
  scale_fill_continuous(high = "#132B43", low = "white") + 
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(legend.position = "none") + 
  geom_text(aes(label = number_of_dd)) +
  scale_x_discrete(position = "top") +
  labs(caption = paste0("Figure 2. The daily double within a category was ",
                        "weighted acrossed all \nmissing values.")) +
  theme(plot.caption = element_text(hjust = 0, size = rel(1.2)))
```

## Results

The heatmaps displaying the percent of daily doubles that occurred at each position is given in Figure 3 (round 1) and Figure 4 (round 2). In round 1, the daily double occured most often (7.33%) in category 1 at a point value of 800. Other points where the daily double occured frequently (> 6%) were category 4 at a point value of 800 (6.81%), and category 1 at a point value of 1000 (6.75%). The least likely placements were across all categories in value 100 (0.35% or less), as well as in categories 2 (4.63% or less) and 6 (4.27% or less). Value 800 is favored across all categories (range: 4.27% – 7.33%). 

In round 2, the daily double was located most often in category 1 at a point value of 1600 (7.06%). The other common positions were category 5, point value 1600 (6.59%) and category 3, point value 1600 (6.41%). The least likely occurrences were across value 100 (0.51% or less) and down categories 2 and 6 (< 4.78% and < 4.75% respectivley). The daily double appeared most frequently at the 1600 point value across all categories (4.75% – 7.06%). Overall, the occurrence of the daily double remained extremely similar across round 1 and round 2 for the years analyzed.

As a secondary question, the heatmaps for all data (rounds 1 and 2) were faceted by year to look at changes over time (Figure 5). Since the rounds were similar in the initial analysis, the data was not separated by round. The results for years 2001 - 2015 look similar to the results by round. The "hot" points are similar as well as the overall pattern. In 2016, there seems to be a shift away from the traditional spots. Previously, category 2 was one of the least likely columns but after 2016 the percentage surpassed the percentages in category 1 (previously, the most likely column). There also seems to be a shift from row 4 to row 3. Before 2016, rows 4 and 5 were the most likely to contain a Daily Double. After 2016, row 3 a higher percentage of Daily Doubles. The pattern after 2016 also seems to have more randomness than prior to 2016.

```{r, warning = FALSE, error = FALSE, message = FALSE, echo= FALSE}
library(readr)
library(plotly)
library(forcats)
library(lubridate)
library(tidyr)
library(purrr)


dd <- read_csv("data/data_final.csv") 
dd <- dd %>% 
  rename("dd_weighting" = "weight") %>% 
  mutate(y_pos = factor(y_pos, levels = c(1, 2, 3, 4, 5)),
         y_pos = factor(y_pos, levels = rev(levels(y_pos))),
         x_pos = factor(x_pos, levels = c(1, 2, 3, 4, 5, 6)),
         x_pos = fct_recode(x_pos, 
                            "Category 1" = "1",
                            "Category 2" = "2",
                            "Category 3" = "3",
                            "Category 4" = "4",
                            "Category 5" = "5",
                            "Category 6" = "6")) 
```

```{r Daily Doubles by Round, echo= FALSE}
dd_1 <- dd %>% 
  filter(round == "1") %>%
  mutate(y_pos = fct_recode(y_pos,
                            "200" = "1",
                            "400" = "2",
                            "600" = "3",
                            "800" = "4",
                            "1000" = "5")) %>% 
  group_by(x_pos, y_pos) %>% 
  summarise(number_of_doubles = sum(dd_weighting)) %>% 
  mutate(Percent = round(number_of_doubles/2476*100, digits = 2)) %>%
  ggplot(aes(x = x_pos, y = y_pos, fill = Percent)) +
  geom_tile(color = "white") +
   labs(x = NULL, 
        y = NULL, 
        fill = "Percent of Daily Doubles") +
  scale_fill_continuous(high = "dodgerblue4", 
                        low = "azure") +
  theme(legend.position = "bottom", 
        legend.box = "horizontal") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black") +
  scale_x_discrete(position = "top")
dd_1
```
\
Figure 3. Percent of Daily Doubles by Position in Round 1
\
\
\
\
```{r echo= FALSE}
dd_2 <- dd %>%
  filter(round == "2") %>%
  mutate(y_pos = fct_recode(y_pos,
                            "400" = "1",
                            "800" = "2",
                            "1200" = "3",
                            "1600" = "4",
                            "2000" = "5")) %>% 
  group_by(x_pos, y_pos) %>% 
  summarise(number_of_doubles = sum(dd_weighting)) %>% 
  mutate(Percent = round(number_of_doubles/4887*100, digits = 2)) %>%
    ggplot(aes(x = x_pos, y = y_pos, fill = Percent)) +
  geom_tile(color = "white") +
   labs(x = NULL, 
        y = NULL, 
        fill = "Percent of Daily Doubles") +
  scale_fill_continuous(high = "dodgerblue4", 
                        low = "azure") +
  theme(legend.position = "bottom", 
        legend.box = "horizontal") +
  geom_text(aes(label = paste0(Percent, "%")), color = "black") +
  scale_x_discrete(position = "top")
dd_2

```
\
Figure 4. Percent of Daily Doubles by Position in Round 2
\
\
\
\
```{r Daily Double Heatmap by year, fig.height = 7, echo= FALSE}
dd %>% 
  mutate(year = year(air_date),
         y_pos = fct_recode(y_pos,
                            "200/400" = "1",
                            "400/800" = "2",
                            "600/1200" = "3",
                            "800/1600" = "4",
                            "1000/2000" = "5")) %>% 
  group_by(x_pos, y_pos, year) %>%
  summarise(number_of_doubles = sum(dd_weighting)) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  mutate(Percent = round(number_of_doubles/sum(number_of_doubles) * 100), 
         digits = 2)  %>% 
  ggplot(aes(x = x_pos, y = y_pos, fill = Percent)) +
  geom_tile(color = "white") +
   labs(x = NULL, 
        y = NULL, 
        fill = "Percent of Daily Doubles") +
  scale_fill_continuous(high = "dodgerblue4", 
                        low = "azure") +
  theme(legend.position = "bottom", 
        legend.box = "horizontal",
        axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(position = "top") +
  facet_wrap(~year, ncol = 4)

```
\
Figure 5. Percent of Daily Doubles by Position Faceted by Year
\
\
\
\

## Conclusions

Our findings  are very similar to previous analyses on Daily Double location by round. One of the original analyses of Daily Double location was conducted by Nathan Yau (https://flowingdata.com/2015/03/03/where-to-find-jeopardy-daily-doubles/) in 2015. The results of this study look similar to his heatmap, which shows the location of all Daily Doubles for 31 seasons. Yau's analysis was done because of a Jeopardy contestant, Arthur Chu, that appeared on the show in 2014. Arthur Chu found success on the show by jumping around the board hunting for the Daily Doubles. He was not the first to adopt this strategy, but was one of the most controversial players because of his approach. Since 2015, there have been other analyses that looked at the same question and found similar results by round. To our knowledge, no other analyses have look at the data by year. This analysis included data up to 2019 and showed a noticeable shift in location around 2015/2016. This shift likely occurred because of Arthur Chu and the subsequent analyses of Daily Double locations. Future analyses may consider evaluating this change over the next few years to see if another pattern is established, or if Jeopardy adopts a more random Daily Double placement.
