---
title: "Final Presentation"
author: "Collin Brehmer, Molly Hischke, Kyle Hancock, Nikki Johnson"
date: "12/13/2019"
output:
  ioslides_presentation: default
  beamer_presentation: default
---

## Research Question

What positions on the Jeopardy board are most likely to be a Daily Double?

## Introduction

- Importance to the game (James Holzhauer)
- Other research
- Challenge of finding position
- Visual representation

## Methods

- Only included episodes from dates beyond November 26, 2001 
- Finding x position of daily double (filtered to episodes with 12 categories)
- Finding y position of daily double (Weighted across missing values)

```{r, echo = FALSE, message = FALSE, include = FALSE}
library(tidyverse)
library(ggplot2)
jeopardy_board <- tibble(
  x_pos = c("Category 1", "Category 1", "Category 1", 
            "Category 1", "Category 1", 
            "Category 2", "Category 2", "Category 2", 
            "Category 2", "Category 2",
            "Category 3", "Category 3", "Category 3", 
            "Category 3", "Category 3", 
            "Category 4", "Category 4", "Category 4", 
            "Category 4", "Category 4", 
            "Category 5", "Category 5", "Category 5", 
            "Category 5", "Category 5",
            "Category 6", "Category 6", "Category 6", 
            "Category 6", "Category 6"),
  y_pos = c(200, 400, 600, 800, 1000, 
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000,
            200, 400, 600, 800, 1000, 
            200, 400, 600, 800, 1000))

jeopardy_board <- jeopardy_board %>% 
  mutate(number_of_dd = c(1, 0, 0, 2, 1,
                          1, 0, 1, 1, 1,
                          1, 1, 1, 1, 1,
                          1, 1, 1, 0, 1,
                          1, 0, 0, 1, 1,
                          1, 1, 1, 0, 1),
          text = c(NA, "MISSING", "MISSING", "Daily Double", NA,
                          NA, "MISSING", NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, "MISSING", NA,
                          NA, "MISSING", "MISSING", NA, NA,
                          NA, NA, NA, "MISSING", NA)) 
weighted_jeopardy_board <- jeopardy_board %>% 
  mutate(number_of_dd = c(NA, 0.33, 0.33, 0.33, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA,
                          NA, NA, NA, NA, NA)) 

```

## Methods
      
```{r, echo = FALSE, fig.height= 4, fig.width = 7, warning = FALSE}
jeopardy_board %>% 
   mutate(y_pos = factor(y_pos), 
                        y_pos = factor(y_pos, levels = rev(levels(y_pos)))) %>% 
  ggplot(aes(x = x_pos, y = y_pos, fill = number_of_dd)) +
  geom_tile(color = "White", size = 0.1) +
  scale_fill_continuous(high = "#132B43", low = "white") + 
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(legend.position = "none") + 
  geom_text(aes(label = text), color = "gray") +
  scale_x_discrete(position = "top") +
  theme(plot.caption = element_text(hjust = 0, size = rel(1.2)))
```

## Methods

```{r, echo = FALSE, fig.height= 4, fig.width = 7, warning = FALSE}
weighted_jeopardy_board %>% 
   mutate(y_pos = factor(y_pos), 
                        y_pos = factor(y_pos, levels = rev(levels(y_pos)))) %>% 
  ggplot(aes(x = x_pos, y = y_pos, fill = number_of_dd)) +
  geom_tile(color = "White", size = 0.1) +
  scale_fill_continuous(high = "#132B43", low = "white") + 
  labs(x = NULL, 
       y = NULL, 
       fill = NULL) +
  theme(legend.position = "none") + 
  geom_text(aes(label = number_of_dd)) +
  scale_x_discrete(position = "top") +
  theme(plot.caption = element_text(hjust = 0, size = rel(1.2)))
```

## Results

Flex Dashboard:

- Daily Double Round 1 since 2001
- Daily Double Round 2 since 2001
- Daily Double Location by year

- Daily Doubles are not evenly distributed across the board
- No differences in daily double location between Round 1 and Round 2
- 15 out of 19 years had daily doubles within the 800/1600 row
- 

## Conclusions

## Overview of Approach

## Interesting Packages or Techniques

- `map()`, `mutate()`, and `join`
- `plot_ly()` 

## Lessons Learned